Predicting Ambient Air Pollution (PM2.5) Across the Contiguous U.S.

### 2023-04-20 Draft.Rmd
boosting

# Check linear regression model performance using cross-validation
folds1 <- vfold_cv(train, v = 10)
folds1
res1 <- fit_resamples(wf1, resamples = folds1)
res1 %>% 
    collect_metrics()
```

@@

# Check performance using cross-validation
folds <- vfold_cv(train, v = 10)
folds
res2 <- fit_resamples(wf2, resamples = folds) # ERROR Assigned data `orig_rows` must be compatible with existing data.
res2 %>% 
    collect_metrics()
```

```{r Bootstrap}
# Create Bootstrap Model (resampling data)
set.seed(123)
model3 <- nearest_neighbor(neighbors = 5) %>% 
    set_engine("kknn") %>% 
    set_mode("classification")

@@

>>>>

# Check linear regression model performance using cross-validation

folds1 <- vfold_cv(train, v = 10)
folds1
res1 <- fit_resamples(wf1, resamples = folds1)
res1 %>% 
    collect_metrics()
```

@@

# Check performance using cross-validation

folds2 <- vfold_cv(train, v = 10)
# folds
res2 <- fit_resamples(wf2, resamples = folds2) # ERROR Assigned data `orig_rows` must be compatible with existing data.
res2 %>% 
    collect_metrics()
```

```{r Bootstrap test}
model3 = boost_tree( mtry = integer(), trees = integer(), min_n = integer(),
  tree_depth = integer(), learn_rate = numeric(), loss_reduction = numeric(),
  sample_size = numeric(), stop_iter = integer()) %>%
  set_engine("xgboost") %>%
  set_mode("regression") %>%
  translate()

# preprocessing "recipe"
preprocessing_recipe <- 
  recipes::recipe(value~lon+lat, data = training(dat_split)) %>%
  # convert categorical variables to factors
  recipes::step_string2factor(all_nominal()) %>%
  # combine low frequency factor levels
  recipes::step_other(all_nominal(), threshold = 0.01) %>%
  # remove no variance predictors which provide no predictive information 
  recipes::step_nzv(all_nominal()) %>%
  prep()

folds3 <- 
  recipes::bake(
    preprocessing_recipe, 
    new_data = training(dat_split)
  ) %>%  
  rsample::vfold_cv(v = 5)

# XGBoost model specification
xgboost_model <- parsnip::boost_tree(mode = "regression", trees = 1000, min_n = tune(),
  tree_depth = tune(), learn_rate = tune(), loss_reduction = tune()) %>%
  set_engine("xgboost", objective = "reg:squarederror")

# grid specification
xgboost_params <- 
  dials::parameters(min_n(), tree_depth(), learn_rate(), loss_reduction())

xgboost_grid <- dials::grid_max_entropy(xgboost_params, size = 60)
knitr::kable(head(xgboost_grid))

xgboost_wf <- 
  workflows::workflow() %>%
  add_model(xgboost_model) %>% 
  add_formula(value~lon+lat)

# hyperparameter tuning
xgboost_tuned <- tune::tune_grid(
  object = xgboost_wf,
  resamples = folds3,
  grid = xgboost_grid,
  metrics = yardstick::metric_set(rmse, rsq, mae),
  control = tune::control_grid(verbose = TRUE)
)

xgboost_tuned %>%
  tune::show_best(metric = "rmse") %>%
  knitr::kable()

```{r Bootstrap}
# Create Bootstrap Model (resampling data)
set.seed(123)
model3 <- boost_tree() %>% 
    set_engine("xgboost") %>%
    set_mode("regression")

@@

# Fit the model on the training dataset using the `fit()` function
res3 <- fit(wf3, data = train) # Error in if (is.numeric(args$trees) && args$trees < 0) { : missing value where TRUE/FALSE needed

# Check performance on the complete training data
res3 %>% 
    extract_fit_engine() %>% 
    summary()

# Check performance using cross-validation

folds3 <- vfold_cv(train, v = 10)
# folds
res3 <- fit_resamples(wf3, resamples = folds3) 
res3 %>% 
    collect_metrics()
```

Preprocessing requirements
xgboost does not have a means to translate factor predictors to grouped splits. Factor/categorical predictors need to be converted to numeric values (e.g., dummy or indicator variables) for this engine. When using the formula method via fit.model_spec(), parsnip will convert factor columns to indicators using a one-hot encoding.
